---
title: "ESM 207 Homework #3"
author: "Pat Byrne"
date: "5/10/2020"
output: html_document
---

```{r setup, include=FALSE}
# Clear global environment
rm(list = ls()) 
# knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(janitor)
library(stargazer)
library(ggthemes)
```


```{r}
# Lock and load in the data
data.df <- read_csv('Homework 3 Data.csv') 
```

```{r}
mwtpl <- data.df$Price_Dollars
mwtph <- data.df$Price_Dollars
ql <- data.df$Q_Low_Gallons
qh <- data.df$Q_High_Gallons
```

```{r}
plot(x = ql, y = mwtpl, xlim = c(0,500000), ylim = c(0,13)) 
plot(x = qh, y = mwtph, xlim = c(0,500000), ylim = c(0,13))
```

```{r}
# In this chunk, I make a data frame for high and low income consumers, use those data frames to fit a linear model for each consumer type, and extract the coefficients of their demand curves
low.df <- tibble(MWTP_L = mwtpl, Q_L = ql)
lm.low <- lm(formula = MWTP_L ~ Q_L, data = low.df)

coef <- as.numeric(lm.low$coefficients) # Spot 1 is the intercept, spot 2 is coef
b_L <- coef[1]
a_L <- coef[2] 


high.df <- tibble(MWTP_H = mwtph, Q_H = qh)
lm.high <- lm(formula = MWTP_H ~ Q_H, data = high.df)

coef <- as.numeric(lm.high$coefficients) # Spot 1 is the intercept, spot 2 is coef
b_H <- coef[1]
a_H <- coef[2]
```

```{r}
# In this chunk I calculate the predicted willingness to pay for a given quantity for each consumer set and then graph them
low.df <- low.df %>% 
  mutate(
    MWTP_predict = b_L + a_L*Q_L
  )

high.df <- high.df %>% 
  mutate(
    MWTP_predict = b_H + a_H*Q_H
  )

ggplot() +
  geom_point(data = low.df, aes(x = Q_L, y = MWTP_L), color = 'red') + 
  geom_line(data=low.df, aes(x = Q_L, y = MWTP_predict), color = 'red', lty = 2) + 
  geom_point(data = high.df, aes(x = Q_H, y = MWTP_H), color = 'blue') + 
  geom_line(data= high.df, aes(x = Q_H, y = MWTP_predict), color = 'blue', lty = 2) +
  theme_tufte()
```

Say we have two linear demand curves of the following form:
For lower income consumers:
$P = b_L + a_L*Q_L$
Then, $Q_L = \frac{P - b_L}{a_L}$
For higher income consumers:
$P = b_H + a_H*Q_H$
Then, $Q_H = \frac{P - b_H}{a_H}$

Aggregate Daily Demand Curve = $Q_T = Q_L + Q_H$
If $Q < \frac{b_L - b_H}{a_H}$  then $P = b_H - a_H*Q$
If $Q >= \frac{b_L - b_H}{a_H}$  then $P = (\frac{a_H * a_L}{a_h + a_L})Q_T + (\frac{a_H}{a_H + a_L})b_L + (\frac{a_L}{a_h + a_L})b_H$

Supply Curve:
$P = (\frac{P_{eq}}{Q_{eq}})Q_T$ where $P_{eq}$ is given and $Q_{eq}$ is found by calculating $Q_H + Q_L$ at $P_{eq}$. We know the intercept is 0, and the slope is just $\frac{rise}{run}$

## Question 1.
```{r}
# From the "following stylized facts," we know that P_eq = $3.00
P_eq = 3
Q_L = (P_eq - b_L)/a_L
Q_H = (P_eq - b_H)/a_H
Q_eq = Q_H + Q_L

# Benefit to producers:
integrand.producers <- function(x) {(P_eq/Q_eq)*x}
integral.producers <- integrate(integrand.producers, 0, Q_eq)
benefit.producers.1 <- (P_eq*Q_eq) - as.numeric(integral.producers$value)

# Benefit to consumers (most easily found by doing each separately and then adding - THIS IS THE ANSWER TO QUESTION 2.):
integrand.low <- function(x) {b_L + a_L*x}
integral.low <- integrate(integrand.low, 0, Q_L)
benefit.low.1 <- as.numeric(integral.low$value) - P_eq*Q_L

integrand.high <- function(x) {b_H + a_H*x}
integral.high <- integrate(integrand.high, 0, Q_H)
benefit.high.1 <- as.numeric(integral.high$value) - P_eq*Q_H

benefit.consumers.1 = benefit.low.1 + benefit.high.1

# Environmental costs: $1.50/gal locally, $0.50/gal globally
env.cost.local.1 = Q_eq*1.5
env.cost.global.1 = Q_eq*0.5

```

## Question 3.
The equation for the supply curve has changed, and is now $P = (\frac{P_{eq}}{Q_{eq}})Q_T + tax$ 
```{r}
# P_eq and Q_eq are the original values. Q_eq.2 is the new equilibrium aggregate quantity of gasoline demanded, and P_eq.2 is the new price per gallon based on that quantity
t = 1 # t = tax
Q_eq.2 = (  (a_H/(a_H + a_L)*b_L) + (a_L/(a_H + a_L)*b_H) - t ) / ( (P_eq/Q_eq) -  (a_H*a_L)/(a_H + a_L)  )
P_eq.2 = (P_eq/Q_eq)*Q_eq.2 + t
Q_L.2 = (P_eq.2 - b_L)/a_L
Q_H.2 = (P_eq.2 - b_H)/a_H


# Benefit to producers:
integrand.producers <- function(x) {(P_eq/Q_eq)*x + t}
integral.producers <- integrate(integrand.producers, 0, Q_eq.2)
benefit.producers.2 <- (P_eq.2*Q_eq.2) - as.numeric(integral.producers$value)

# Benefit to consumers (most easily found by doing each separately and then adding - THIS IS THE ANSWER TO QUESTION 2.):
integrand.low <- function(x) {b_L + a_L*x}
integral.low <- integrate(integrand.low, 0, Q_L.2)
benefit.low.2 <- as.numeric(integral.low$value) - P_eq.2*Q_L.2

integrand.high <- function(x) {b_H + a_H*x}
integral.high <- integrate(integrand.high, 0, Q_H.2)
benefit.high.2 <- as.numeric(integral.high$value) - P_eq.2*Q_H.2

benefit.consumers.2 = benefit.low.2 + benefit.high.2

# Environmental costs: $1.50/gal locally, $0.50/gal globally
env.cost.local.2 = Q_eq.2*1.5
env.cost.global.2 = Q_eq.2*0.5

tax_revenue.1 = Q_eq.2*t

```

## Question 4.
```{r}
# Here's where we get funky. We're going to rerun the previous chunk multiple times according to the stipulations of question 4, with a different value of the tax each time

consumption.proportion.low = Q_L/Q_eq
p.low <- consumption.proportion.low

consumption.proportion.high = Q_H/Q_eq
p.high <- consumption.proportion.high


taxes <- seq(0,5, by = 0.5)
gas.produced <- c()
gas.price <- c()
welfare.low <- c()
welfare.high <- c()
welfare.producers <- c()

for (i in 1:length(taxes)) {
  Q.eq = (  (a_H/(a_H + a_L)*b_L) + (a_L/(a_H + a_L)*b_H) - taxes[i] ) / ( (P_eq/Q_eq) -  (a_H*a_L)/(a_H + a_L)  )
  P.eq = (P_eq/Q_eq)*Q.eq + taxes[i]
  
  gas.produced[i] <- Q.eq
  gas.price[i] <- P.eq
  
  Q.L = (P.eq - b_L)/a_L
  Q.H = (P.eq - b_H)/a_H
  
  # Producer welfare
  integrand.producers <- function(x) {(P_eq/Q_eq)*x + taxes[i]}
  integral.producers <- integrate(integrand.producers, 0, Q.eq)
  welfare.producers[i] <- (P.eq*Q.eq) - as.numeric(integral.producers$value)
  
  # Tax revenue
  tax_revenue = Q.eq*taxes[i]
  
  # Local environmental damage
  local_damage = Q.eq*1.5
  
  # High income welfare
  integrand.high <- function(x) {b_H + a_H*x}
  integral.high <- integrate(integrand.high, 0, Q.H)
  welfare.high[i] <- as.numeric(integral.high$value) - P.eq*Q.H + tax_revenue*p.high
  
  # Low income welfare
  integrand.low <- function(x) {b_L + a_L*x}
  integral.low <- integrate(integrand.low, 0, Q.L)
  welfare.low[i] <- as.numeric(integral.low$value) - P.eq*Q.L + tax_revenue*p.low - local_damage
}

results.question.4 <- tibble(taxes, gas.produced, gas.price, welfare.producers, welfare.high, welfare.low)

```

## Question 5.
```{r}
# The difference between this question and question 1. is that b_L and b_H are both divided by two
Q_eq.5 = (  (a_H/(a_H + a_L)*(b_L*0.5)) + (a_L/(a_H + a_L)*(b_H*0.5)) ) / ( (P_eq/Q_eq) -  (a_H*a_L)/(a_H + a_L)  )
P_eq.5 = (P_eq/Q_eq)*Q_eq.5 # Gas Price
Q_L.5 = (P_eq.5 - b_L*0.5)/a_L # Low income consumption
Q_H.5 = (P_eq.5 - b_H*0.5)/a_H # High income consumption

# Compare these two numbers to those found in question 1.
env.cost.local.5 = Q_eq.5*1.5
env.cost.global.5 = Q_eq.5*0.5


```

## Question 6.
```{r}
# We iterate this chunk with different values of t until env.cost.local.6 - env.cost.local.5 is 0
t.6 = 7.29985
Q_eq.6 = (  (a_H/(a_H + a_L)*b_L) + (a_L/(a_H + a_L)*b_H) - t.6 ) / ( (P_eq/Q_eq) -  (a_H*a_L)/(a_H + a_L)  )

env.cost.local.6 = Q_eq.6*1.5

local.damage.diff = env.cost.local.6 - env.cost.local.5
```


## Summary Chunk:
Run this chunk and then open up the console
```{r}
# Question 1.
# Aggregate daily demand curve is above the Question 1. chunk, as is the supply curve
# Benefit to consumers
benefit.consumers.1
# Benefit to producers
benefit.producers.1
# Environmental Costs
# Locally
env.cost.local.1
# Globally
env.cost.global.1 

# Question 2.
benefit.low.1 
benefit.high.1

# Question 3.
# a.
Q_eq - Q_eq.2
Q_L - Q_L.2
Q_H - Q_H.2
# b.
P_eq - P_eq.2
# c.
benefit.high.1 - benefit.high.2
# d.
benefit.low.1 - benefit.low.2
# e.
benefit.producers.1 - benefit.producers.2
# f.
env.cost.local.1 - env.cost.local.2
# g.
env.cost.global.1 - env.cost.global.2
# h. 
tax_revenue.1

# Question 4.
# a. b. c.
results.question.4

# Question 5.
# a.
Q_H - Q_H.5  
# b.
Q_L - Q_L.5 
# c.
P_eq - P_eq.5 
# d.
env.cost.local.1 - env.cost.local.5
# e.
env.cost.global.1 - env.cost.global.5

# Question 6. 
t.6
env.cost.local.5 - env.cost.local.6 # .5 is electric cars and .6 is $T/gal tax  -  in the code, T = t.6
```













